Toward \earlywarning\ detection of gravitational waves from compact binary coalescence

### 2012-01-06 article.tex
Correction to appendix: above, not below
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1449 ccdd05d5-56c1-461a-bf98-85f6876523b4

The expected matched-filter SNR, integrated from $f_\mathrm{low}$ to $f_\mathrm{high}$, is given by equation~(\ref{eq:expected-snr}).  The high frequency cutoff for the inspiral is frequently taken to be the gravitational wave frequency at the ISCO; for non-spinning systems, $f_\mathrm{ISCO} = 4400 (\Msun / M)$~Hz, where $M$ is the total mass of the binary.  The choice of $f_\mathrm{low}$ is based on the fraction of the total SNR that is accumulated at frequencies above $f_\mathrm{low}$.  To illustrate the relative contributions to the SNR at different frequencies for a (1.4,~1.4)~$\Msun$ binary, we normalized and plotted the integrand of equation~(\ref{eq:expected-snr}), the noise-weighted power spectral density of the inspiral waveform, in Figure~\ref{fig:low-frequency-cutoff}b.  This is the quantity
%
$$
	\frac{1}{\rho^2}\frac{\mathrm{d}\rho^2}{\mathrm{d}f} = \frac{f^{-7/3}}{S(f)} \left( \int_0^{f_\mathrm{ISCO}} \frac{{f'}^{-7/3}}{S(f')} \, \mathrm{d}f' \right)^{-1},
$$
%
which is normalized by the total SNR squared in order to put detectors with different absolute sensitivities on the same footing.  We used several different noise power spectra: all of the envisioned Advanced LIGO configurations from~\citet{ALIGONoise}; the best achieved sensitivity at LIGO Hanford Observatory (LHO) in LIGO's fifth science run (S5), measured by \citet{S5InspiralRange}; and the best achieved sensitivity at LHO during S6, measured by \citet{S6InspiralRange}.  (The noise spectra themselves are shown in Figure~\ref{fig:low-frequency-cutoff}a.)  It is plain that during S5 and S6 the greatest contribution to the SNR was between 100 and 150~Hz, but for all of the proposed Advanced LIGO configurations the bulk of the SNR is accumulated below 60~Hz.  This information is presented in a complementary way in Figure~\ref{fig:low-frequency-cutoff}c, as the square root of the cumulative integral from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, interpreted as a fraction of the total ``available'' SNR,
%
$$
	\rho_\mathrm{frac}(f_\mathrm{low}) = \sqrt{\left( \int_{f_\mathrm{low}}^{f_\mathrm{ISCO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right) \left( \int_0^{f_\mathrm{ISCO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right)^{-1}}.
$$
%
Table~\ref{table:accum_snr} shows the fractional accumulated SNR for four selected low frequency cutoffs, 40~Hz, 30~Hz, 20~Hz, and 10~Hz.  In S5 and S6, all of the SNR is accumulated above 40~Hz.  For the `high frequency' Advanced LIGO configuration, almost 50\% of the SNR is accumulated \emph{below} 40~Hz.  For the preferred final configuration, 'zero detuning, high power,' 86.1\% of the SNR is below 40~Hz, 93.2\% is below 30~Hz, and 98.1\% is below 20~Hz.

>>>>

The expected matched-filter SNR, integrated from $f_\mathrm{low}$ to $f_\mathrm{high}$, is given by equation~(\ref{eq:expected-snr}).  The high frequency cutoff for the inspiral is frequently taken to be the gravitational wave frequency at the ISCO; for non-spinning systems, $f_\mathrm{ISCO} = 4400 (\Msun / M)$~Hz, where $M$ is the total mass of the binary.  The choice of $f_\mathrm{low}$ is based on the fraction of the total SNR that is accumulated at frequencies above $f_\mathrm{low}$.  To illustrate the relative contributions to the SNR at different frequencies for a (1.4,~1.4)~$\Msun$ binary, we normalized and plotted the integrand of equation~(\ref{eq:expected-snr}), the noise-weighted power spectral density of the inspiral waveform, in Figure~\ref{fig:low-frequency-cutoff}b.  This is the quantity
%
$$
	\frac{1}{\rho^2}\frac{\mathrm{d}\rho^2}{\mathrm{d}f} = \frac{f^{-7/3}}{S(f)} \left( \int_0^{f_\mathrm{ISCO}} \frac{{f'}^{-7/3}}{S(f')} \, \mathrm{d}f' \right)^{-1},
$$
%
which is normalized by the total SNR squared in order to put detectors with different absolute sensitivities on the same footing.  We used several different noise power spectra: all of the envisioned Advanced LIGO configurations from~\citet{ALIGONoise}; the best achieved sensitivity at LIGO Hanford Observatory (LHO) in LIGO's fifth science run (S5), measured by \citet{S5InspiralRange}; and the best achieved sensitivity at LHO during S6, measured by \citet{S6InspiralRange}.  (The noise spectra themselves are shown in Figure~\ref{fig:low-frequency-cutoff}a.)  It is plain that during S5 and S6 the greatest contribution to the SNR was between 100 and 150~Hz, but for all of the proposed Advanced LIGO configurations the bulk of the SNR is accumulated below 60~Hz.  This information is presented in a complementary way in Figure~\ref{fig:low-frequency-cutoff}c, as the square root of the cumulative integral from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, interpreted as a fraction of the total ``available'' SNR,
%
$$
	\rho_\mathrm{frac}(f_\mathrm{low}) = \sqrt{\left( \int_{f_\mathrm{low}}^{f_\mathrm{ISCO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right) \left( \int_0^{f_\mathrm{ISCO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right)^{-1}}.
$$
%
Table~\ref{table:accum_snr} shows the fractional accumulated SNR for four selected low frequency cutoffs, 40~Hz, 30~Hz, 20~Hz, and 10~Hz.  In S5 and S6, all of the SNR is accumulated above 40~Hz.  For the `high frequency' Advanced LIGO configuration, scarcely \emph{half} of the SNR is accumulated above 40~Hz.  For the preferred final configuration, 'zero detuning, high power,' 86.1\% of the SNR is above 40~Hz, 93.2\% is above 30~Hz, and 98.1\% is above 20~Hz.
